---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  html_document: default
  pdf_document: default
  word_document: default
---
<center> <h1> Assignment 1 (10%) </h1> </center>
<center>  <h3> [Insert your full name] </h2> </center>
<center> <h3> [Insert course section & student number] </h2> </center>
---

#### 1. Read the csv files in the folder. (4 points)
```{r}

macro<-read.csv("C:/Users/Marco/Documents/CMTH 642/Assignment 1/USDA_Macronutrients.csv",header= T)
micro<-read.csv("C:/Users/Marco/Documents/CMTH 642/Assignment 1/USDA_Micronutrients.csv",header= T)

```

#### 2. Merge the data frames using the variable "ID". Name the Merged Data Frame "USDA". (4 points)
```{r}
df<-merge(macro,micro, by = "ID")
```

#### 3. Check the datatypes of the attributes. Delete the commas in the Sodium and Potasium records. Assign Sodium and Potasium as numeric data types. (4 points)
```{r}
df$Sodium = gsub(",", "", df$Sodium)
df$Potassium = gsub(",", "", df$Potassium)
df$Sodium = as.numeric(df$Sodium)
df$Potassium = as.numeric(df$Potassium)
```

#### 4. Remove records (rows) with missing values in more than 4 attributes (columns). How many records remain in the data frame? (4 points)
```{r}
summary(df)

CountNA=apply(is.na(df),1,sum)
df=df[CountNA<5,]
nrow(df) #remaining rows 

```

#### 5. For records with missing values for Sugar, Vitamin E and Vitamin D, replace missing values with mean value for the respective variable. (4 points)
```{r}
df$Sugar[is.na(df$Sugar)] = mean(df$Sugar[!is.na(df$Sugar)])
df$VitaminE[is.na(df$VitaminE)] = mean(df$VitaminE[!is.na(df$VitaminE)])
df$VitaminD[is.na(df$VitaminD)] = mean(df$VitaminD[!is.na(df$VitaminD)])
```

#### 6. With a single line of code, remove all remaining records with missing values. Name the new Data Frame "USDAclean". How many records remain in the data frame? (5 points)
```{r}
USDAclean=df[complete.cases(df),]
nrow(USDAclean)
```

#### 7. Which food has the highest sodium level? (5 points)
```{r}
USDAclean$Description[which.max(USDAclean$Sodium)]

```

#### 8. Create a histogram of Vitamin C distribution in foods. (5 points)
```{r}

hist(USDAclean$VitaminC, ylim=range(1,100),xlab = paste("Vitamin C"),main = paste("Vitamin C Distribution"))
```

#### 9. Create a boxplot to illustrate the distribution of values for TotalFat, Protein and Carbohydrate. (5 points)
```{r}
boxplot(USDAclean[4:6])
```

#### 10. Create a scatterplot to illustrate the relationship between a food's TotalFat content and its Calorie content. (5 points)
```{r}
library(ggvis)
USDAclean %>% ggvis(x = ~TotalFat, y = ~Calories) %>% layer_points()
```

#### 11. Add a variable to the data frame that takes value 1 if the food has higher sodium than average, 0 otherwise. Call this variable HighSodium. Do the same for High Calories, High Protein, High Sugar, and High Fat. How many foods have both high sodium and high fat? (5 points)
```{r}

USDAclean$HighSodium[USDAclean$Sodium > mean(USDAclean$Sodium)] <- 1
USDAclean$HighSodium[USDAclean$Sodium <= mean(USDAclean$Sodium)] <- 0

USDAclean$HighCalories[USDAclean$Calories > mean(USDAclean$Calories)] <- 1
USDAclean$HighCalories[USDAclean$Calories <= mean(USDAclean$Calories)] <- 0

USDAclean$HighProtein[USDAclean$Protein > mean(USDAclean$Protein)] <- 1
USDAclean$HighProtein[USDAclean$Protein <= mean(USDAclean$Protein)] <- 0

USDAclean$HighSugar[USDAclean$Sugar > mean(USDAclean$Sugar)] <- 1
USDAclean$HighSugar[USDAclean$Sugar <= mean(USDAclean$Sugar)] <- 0

USDAclean$HighFat[USDAclean$TotalFat > mean(USDAclean$TotalFat)] <- 1
USDAclean$HighFat[USDAclean$TotalFat <= mean(USDAclean$TotalFat)] <- 0


cat(sum(apply(USDAclean[c("HighSodium", "HighFat")], 1, function(x) sum(x) == 2)))
```

#### 12. Calculate the average amount of iron, for high and low protein foods. (5 points)
```{r}
MeanProteinIron <- aggregate(USDAclean$Iron,list(USDAclean$HighProtein),FUN = mean) 
colnames(MeanProteinIron) <- c("low/high protein","AVG")
head(MeanProteinIron)
```

#### 13. Create a function for a "HealthCheck" program to detect unhealthy foods. Use the algorithm flowchart below as a basis. (5 points)
```{r}
require(jpeg)
img<-readJPEG("HealthCheck.jpg")
plot(1:4, ty = 'n', ann = F, xaxt = 'n', yaxt = 'n')
rasterImage(img,1,1,4,4)
```
```{r}
HealthCheck <- function (sodium,sugar,fat) {
  ifelse (sodium==0, "Pass", ifelse (sugar==0, "Pass", ifelse (fat==0, "Pass", "Fail")))
}
```

#### 14. Add a new variable called HealthCheck to the data frame using the output of the function. (5 points)
```{r}
USDAclean$HealthCheck = HealthCheck(USDAclean$HighSodium, USDAclean$HighSugar, USDAclean$HighFat)
```

#### 15. How many foods in the USDAclean data frame fail the HealthCheck? (5 points)
```{r}
sum(USDAclean$HealthCheck == "Fail",na.rm = TRUE)
```

#### 16. Visualize the correlation among Calories, Protein, Total Fat, Carbohydrate, Sodium and Cholesterol. (5 points)
```{r}
cor(USDAclean[3:8])
```

#### 17. Is the correlation between Calories and Total Fat statistically significant? Why? (5 points)
```{r}

cor.test(USDAclean$Calories,USDAclean$TotalFat)
#yes the corelation between calories and Total fat is 0.80449 making it statistically significant due to its linear relationship. Also the p value is less than 0.05 making it statistically significant. The relationship would be described as a high positive relationship. 
```

#### 18. Create a Linear Regression Model, using Calories as the dependent variable Protein, Total Fat, Carbohydrate, Sodium and Cholesterol as the independent variables. (5 points)
```{r}
MOD=summary(lm(Calories~Protein+TotalFat+Carbohydrate+Sodium+Cholesterol,data=USDAclean))
print(MOD)
```

#### 19. Which independent variable is the least significant? Why? (Hint: Use ANOVA) (5 points)
```{r}

summary(aov(MOD,data = USDAclean))

#ANSWER: Sodium is the least significant variable since the F value is the highest making it less significant in comparison to the other variables. Not only that but the P value is above 0.05 as well which would also make it not statistically significant. 

```
ANSWER: Sodium is the lest significant variable since the F value is the highest making it less significant in comparison to the other variables.

#### 20. A new product is just produced with the following data: Protein=0.1, TotalFat=37, Carbohydrate=400, Sodium=440, Cholesterol=75, Sugar=NA, Calcium=35, Iron=NA, Potassium=35, VitaminC=10,	VitaminE=NA, VitaminD=NA. Based on the model you created, what is the predicted value for Calories? (5 points)
```{r}
lm(Calories~Protein+TotalFat+Carbohydrate+Sodium+Cholesterol,data=USDAclean)

pred_value=3.9882753+(0.1)*3.9891994+(37)*8.7716980+(400)*3.7432001+(440)*0.0003383+(75)*0.0110138
print(pred_value)
#The predicted value would be 1827.195
```

#### 21. If the Sodium amount increases from 440 to 44440 (10000% increase), how much change will occur on Calories in percent? Explain why? (5 points)
```{r}
pred_value=3.9882753+(0.1)*3.9891994+(37)*8.7716980+(400)*3.7432001+(44440)*0.0003383+(75)*0.0110138
print(pred_value)
increase=1842.08-1827.195
percentIncrease=increase/1827.195*100
print(percentIncrease)
# due to the low significance of the sodium coefficient any change will have a small change in the predictive value. This is why the percent increase is only 0.81 percent.
```


This is the end of Assignment 1

Ceni Babaoglu, PhD
